---
title: "Intermediate Elective"
date: "05/02/2026"
author: "Sami Gibson"
format: pdf
editor: source
execute: 
  message: false
  warning: false
---
# Set up

## Packages
```{r}
#| label: packages

# reads in tidyverse package from library 
library(tidyverse)
# reads in janitor package from library 
library(janitor)
# reads in here package from library 
library(here)
# reads in naniar package from library 
library(naniar)
# reads in patchwork package from library 
library(patchwork)
# reads in viridis package from library 
library(viridis)
# reads in scales package from library 
library(scales)
```
## Data
```{r}
#| label: data
# read in weather data from NOAA weather station
weather <- read_csv(here("data", "NOAA-weather-data.csv"))
# read in water quality survey metadata
metadata <- read_csv(here("data", "NCOS_YSI_Water_Quality_Monitoring_0.csv"))
# read in water quality parameter measurements
water_parameters <- read_csv(here("data", "YSI_Data_Begin_1.csv"))
# reads in eBird data as birds
birds <- read_csv(here("data", "birds.csv"))
# read in Venco water data
venoco_water <- read_csv(here("data","venoco-water.csv"))
```

# 1. Explain your inspiration
:::{.callout-note}
In 1-2 sentences each, describe:

which visualization you chose for your inspiration
why it makes sense for your dataset of choice
which variables from your dataset will be shown in your visualization, and what visual components they map onto (axes, shapes, colors, etc.)
:::
**Which visualization you chose for your inspiration**: For my inspiration, I chose Steven Ponce's ["Coal falls. Renewables rise."](https://stevenponce.netlify.app/data_visualizations/30DayChartChallenge/2026/30dcc_2026_24.html) visualization because it uses a clean time-series design to show change over time which was visually appealing to me. I am applying that same idea to bird survey data by showing how observed bird counts change across water years. 

**Why it makes sense for your dataset of choice**: This inspiration makes sense for my dataset because the observations in the Ebird data are naturally time-based so a time-series figure can show whether selected species increased, decreased, or stayed stable over time. Instead of showing energy sources over time (like Steven Ponce's origional does), my figure will show selected bird species observed at NCOS over time.

**Which variables from your dataset will be shown in your visualization, and what visual components they map onto (axes, shapes, colors, etc.)**: The variables in shown in my figure will be `water_year` (found from origional `date` column), `species`, and `count` (summed to `total_observations`). Water year will be on the x-axis, total observations will be on the y-axis, species will determine line and point color color, and there will be labels to identify each species at the end of its line.


# 2. Plan your figure
:::{.callout-note}
Sketch what your figure would look like following your inspiration.

Insert that sketch (as a photo, screenshot, or otherwise) into your document.

:::
XX ADD PIC
```{r}
#| label: figure plan image

```

# 3. Code your figure
:::{.callout-note}
Clean/wrangle your data as needed.

Then, code the figure you decided to create.

Make sure all your code is annotated, and that the output is showing.

At the end of this section, include code to save your final figure as a .jpg or .png file.
:::
```{r}
# creates new object birds_clean from birds that has standardized column names
birds_clean <- birds |> clean_names()
```

# 4. Describe your process
:::{.callout-note}
In 1-3 sentences each, describe:

your process of using someone else’s code to create a visualization
challenges you encountered as you made your visualization
how your visualization differs from visualizations you made in class, for assignments, or for a previous elective
:::